Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-19 paper/icml_drau/advup.tex

% \LB{In next equation, tiny detail but I would write $\frac{1}{\deltat}(r\deltat+...)$, easier to read.}
% \begin{align}
% 	A^\pi_\deltat(s, a)=&\reward(s, a)+\frac{\gamma^{\deltat}\E_{\delta s \sim \delta s(s, a)}V^\pi_\deltat(s + \delta s)-V^\pi_\deltat(s)}{\deltat}\nonumber\\
% 	=&\reward(s, a)+\log(\gamma) V^\pi(s)+\partial_s V^\pi(s)f(s, a)\nonumber\\
%          &+\frac{\text{tr}\left(\Sigma^T(s, a)\partial^2_{s^2}V^\pi(s) \Sigma(s,a)\right)}{2} + O(\deltat)\nonumber.
% \end{align}

@@

\subsection{Scaling the Learning Rates}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the discrete algorithm to admit a continuous time limit, the trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.  Intuitively, this means that optimization steps should not be too large,
in which case parameter trajectories could diverge in a single gradient step,
or too small, in which case the trajectories would converge to single points as
$\deltat$ goes to $0$.  This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the tine discretization.

The updates equations on $\theta$ and $\psi$ are described in
Eq.~\eqref{eq:approx_bellman_A}.  To have parameter steps that do not explode or shrink to $0$ too
fast as $\deltat \rightarrow 0$, both $\delta \theta$ and
$\delta \psi$ should verify stochastic differential equations, and notably
have a deterministic component of order $\deltat$. As $\deltat$ goes to $0$,
updates are rephrased as
\begin{align}
	\delta Q_\deltat &= A(s, a)\deltat - r\deltat - \log(\gamma) V(s)\deltat - \partial_s V(s) \delta s\nonumber \\&- \frac{(\delta s)^T \partial_{s^2}^2 V(s) \delta s}{2} + \smallo\left(\deltat\right)\\
	\delta \theta_\deltat &= \eta^V_\deltat \partial_{\theta} V(s) \frac{\delta Q_\deltat}{\deltat} + \smallo\left(\eta^V_\deltat\right)\\
	\delta \psi_\deltat &= \eta^A_\deltat \partial_{\psi} A(s) \frac{\delta Q_\deltat}{\deltat} + \smallo\left(\eta^A_\deltat\right).
\end{align}

For parameter updates to follow a SDE, one must
have $\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$. More precisely
\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$.
		\item If $\beta > 1$, parameter trajectories shrink to a single point as 
			$\deltat$ goes to $0$. \TODO{Is that clear}
	\end{itemize}
	\label{th:cont-params}
\end{theorem}

>>>>

% \LB{In next equation, tiny detail but I would write $\frac{1}{\deltat}(r\deltat+...)$, easier to read.}
% \begin{align}
% 	A^\pi_\deltat(s, a)=&\reward(s, a)+\frac{\gamma^{\deltat}\E_{\deltas \sim \deltas(s, a)}V^\pi_\deltat(s + \deltas)-V^\pi_\deltat(s)}{\deltat}\nonumber\\
% 	=&\reward(s, a)+\log(\gamma) V^\pi(s)+\partial_s V^\pi(s)f(s, a)\nonumber\\
%          &+\frac{\text{tr}\left(\Sigma^T(s, a)\partial^2_{s^2}V^\pi(s) \Sigma(s,a)\right)}{2} + O(\deltat)\nonumber.
% \end{align}

@@

\subsection{Scaling the Learning Rates}
\label{subsec:lr}
\input{pendulum_fig.tex}
For the discrete algorithm to admit a continuous time limit, the trajectories
of parameters must converge to well defined trajectories as $\deltat$ goes to
$0$.  The optimization steps should not be too large,
in which case parameter trajectories could jump in a single gradient step,
or too small, in which case the trajectories would become stationary as
$\deltat$ goes to $0$.  This in turn imposes conditions on the scalings of the
learning rates of algorithms robust to changes of the tine discretization.

The updates of $\theta$ and $\psi$ are described in
\eqref{eq:approx_bellman_A}.  To have parameter updates that do not explode or shrink to $0$ too
fast as $\deltat \rightarrow 0$, both $\delta \theta$ and
$\delta \psi$ should verify stochastic differential equations in the
limit, and notably,
have a deterministic component of order $\deltat$. As $\deltat$ goes to $0$,
a Taylor expansion yields (with $\deltas\deq s'-s$)
\begin{align}
	\delta Q_\deltat &= A(s, a)\,\deltat - r\,\deltat - \log(\gamma)
	\,V(s)\,\deltat - \partial_s V(s) \,\deltas\nonumber \\&- \frac{(\deltas)^T \partial_{s}^2 V(s) \deltas}{2} + \smallo\left(\deltat\right)\\
	\delta \theta_\deltat &= \eta^V_\deltat \partial_{\theta} V(s)
	\frac{\delta Q_\deltat}{\deltat} \\%+ \smallo\left(\eta^V_\deltat\right)\\
	\delta \psi_\deltat &= \eta^A_\deltat \partial_{\psi} A(s)
	\frac{\delta Q_\deltat}{\deltat} %+ \smallo\left(\eta^A_\deltat\right).
\end{align}

If we set 
$\eta^V_\deltat = \alpha^V \deltat$ and $\eta^A_\deltat = \alpha^A
\deltat$, the parameter updates will follow a Euler discretization scheme
for a continuous-time equation, thus presumably converging. This is
actually the only possibility; more precisely,

\begin{theorem}
	\TODO{We need a bit more details: we probably want to assume that states and actions are
	drawn from a single fixed trajectory, and that we are learning for a fixed, constant policy.}
	Let $\eta^V_\deltat = \alpha^V \deltat^\beta$ (resp. $\eta^A \deltat^\beta$) be the learning rate
	of $V$ (resp. $A$), when training on a single exploratory trajectory. Then
	\begin{itemize}
		\item If $\beta = 1$ the discrete parameter trajectories converge to continuous parameter
			trajectories as $\deltat$ goes to $0$.
		\item If $\beta < 1$, parameters grow arbitrarily large in a single timestep as $\deltat$
			goes to $0$.
		\item If $\beta > 1$, parameter trajectories shrink to a single point as 
			$\deltat$ goes to $0$. \TODO{Is that clear}
	\end{itemize}
	\label{th:cont-params}
\end{theorem}
